---
title: "Extract data from ecosystem models for the Southern Ocean"
author: "Camilla Novaglio"
date: "27/04/2023"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### AIM

Extract mean size-resolved total consumer biomass for the period 1990-1999 across FishMIP models 

### Data source 

Size-resolved biomass (tcblog10) is as for EmergentConstrint project. Adding macroecological for this project: 

Download on 01/05/2023 - only history (and picontrol as in same folder) and not future as this project does not require it

scp -r b381217@levante.dkrz.de:/work/bb0820/ISIMIP/ISIMIP3b/OutputData/marine-fishery_global/MACROECOLOGICAL/gfdl-esm4/historical/*nat_default_tcblog10* ./ 

scp -r b381217@levante.dkrz.de:/work/bb0820/ISIMIP/ISIMIP3b/OutputData/marine-fishery_global/MACROECOLOGICAL/ipsl-cm6a-lr/historical/*nat_default_tcblog10* ./ 

### SO boundaries 

From Paul Spence: 
/home/ubuntu/extract_SouthernOcean/Input/1deg_om2.ocean_grid.nc" OR new regridded version but need to ask Paul again. 

LME from - not much data included - from Julia's dropbox folder FishMIP Repos/LME66/

World_Seas_IHO_v3 from https://www.marineregions.org/downloads.php

Or just latitudinal limits to mimic Yang et al. 2022 **used this**

### *TO CHECK* 

Extent is different for dbpm and zoom IPSL - now dealt with it but **needs more checking**. Also need to adjust the emergent constraint code accordingly (e.g. inside plot_gridded_slope_mpas() in plotting_functions_EC you are using setExtent too). 

**Check better** extent, crs, crop and mask inside function and in general 

Final values are mean g m-2 for the period 1990-1999 -> for models providing monthly values these are 1) annual means and 2) period means; for models providing annual values these are period means. These outputs have no time dimension (i.e. they represent a stock and not a rate, as opposed to NPP which is in s-1 or to catches which should be in e.g. year-1). Hence, here to get annual values we average instead of sum across months. For NPP and catch we would sum across months to get yearly value as these are rates with a time dimension.      

To transform from g m-2 to g C m-2 for better comparison with inputs I did /10   

MACROECOLOGICAL: 5 bins, data is extracted but not sure whether I should include it in the multi-model mean as I don't know which classes these are (likely 1 to 5 though as specified in bins names). **Ryan will check**

### Set environment 

```{r environment, echo = FALSE}

rm(list=ls())

library(ncdf4) 
library(raster) 
library(ggplot2) 
library(data.table)
library(dtplyr)
library(ncdf4.helpers)
library(tictoc)
library(dplyr, warn.conflicts = FALSE)
library(broom)
library(tibble)
library(tidyr)
library(purrr)
library(stringr)
library(parallel)
library(pbapply)

select<-dplyr::select
summarise<-dplyr::summarise

# load functions 
source("/home/ubuntu/Extract_SouthernOcean/R/Helper_functions_SO.r")

```

### Read in files and extract modelled data at global scale

```{r step 1 to 2}

# create file names ----
dir<-"/rd/gem/private/users/camillan/EmergentConstraintData" 

mem<-c("apecosm","boats", "dbpm", "zoomss", "ecotroph", "macroecological") 
esm<-c("gfdl-esm4", "ipsl-cm6a-lr")
scenario<-c("historical") # , "ssp126", "ssp585") # NO NEED FOR FUTURE ... 

# all 
combinations<-expand.grid(mem = mem, esm = esm, scenario = scenario) %>%
  mutate(resolution = ifelse(mem %in% c("zoomss", "ecotroph", "macroecological"), "annual", "monthly"), 
         year = case_when(
           mem %in% c("boats", "zoomss", "ecotroph", "macroecological") & scenario == "historical" ~ "1950_2014", 
           mem %in% c("apecosm", "dbpm") & scenario == "historical" ~ "1850_2014",
           scenario %in% c("ssp126", "ssp585") ~ "2015_2100"), 
         mem = as.character(mem), 
         esm = as.character(esm), 
         scenario = as.character(scenario),
         netcdf_name = paste0(paste(mem, esm, "nobasd", scenario, "nat_default_tcblog10_global", resolution, year, sep ="_"), ".nc"),
         esm_simpler = ifelse(esm == "ipsl-cm6a-lr", "ipsl","gfdl"),
         identifier = paste(mem, esm_simpler, scenario, sep ="_")) %>% 
  select(-esm_simpler) %>% 
  arrange(mem, esm, scenario)

# apply function in // ----

netcdf = combinations$netcdf_name

brick_data_annual_subset_mean_ant<-mclapply(netcdf, function(x) extract_antarctica(x, file = "new"), mc.cores = detectCores()-2)
names(brick_data_annual_subset_mean_ant) <- combinations$identifier

### WARNINGs
# land between GFDL and IPSL is different so coastal values might be "biased" e.g. mean with na.rm=TRUE will consider only GFDL values for the very coastal gridcell (and only a set of models).
# ZOOM bins fixed but need to contact Jase and Matthias 

```

### save data

```{r}

# codes to explore some differences between appraoches (all sorted): 
# month = annual /12 
# WS = World_Seas_IHO_v3
# 5 = including macroecological 
# 60 = -60 lat
# checks done with g m-2 values. 

save(brick_data_annual_subset_mean_ant, file = "/home/ubuntu/Extract_SouthernOcean/Tmp_data/01_extract_modelled_data_SO.RData")

```

### multimodel means 

```{r}

# rm(list=ls())
# load("/home/ubuntu/Extract_SouthernOcean/Tmp_data/01_extract_modelled_data_SO.RData")

# do size classes 2:5 that are in common to all models 

mean <-list()

for(i in 2:5){ # note - this becomes 1:5 for WS option (or any other polygon). 

  trial<-stack(
    brick_data_annual_subset_mean_ant$apecosm_ipsl_historical[[i]],
    brick_data_annual_subset_mean_ant$boats_gfdl_historical[[i]], 
    brick_data_annual_subset_mean_ant$boats_ipsl_historical[[i]],
    brick_data_annual_subset_mean_ant$dbpm_ipsl_historical[[i]], 
    brick_data_annual_subset_mean_ant$ecotroph_gfdl_historical[[i]],
    brick_data_annual_subset_mean_ant$ecotroph_ipsl_historical[[i]],
    brick_data_annual_subset_mean_ant$macroecological_gfdl_historical[[i]],
    brick_data_annual_subset_mean_ant$macroecological_ipsl_historical[[i]],
    brick_data_annual_subset_mean_ant$zoomss_gfdl_historical[[i]], 
    brick_data_annual_subset_mean_ant$zoomss_ipsl_historical[[i]])
 
  mean[[i]] <- stackApply(trial, indices =  rep(1,nlayers(trial)), fun = "mean", na.rm = T)
   
}

# do size class 1 where macroecological is available but not boats

trial_1<-stack(
    brick_data_annual_subset_mean_ant$apecosm_ipsl_historical[[1]],
    brick_data_annual_subset_mean_ant$dbpm_ipsl_historical[[1]],
    brick_data_annual_subset_mean_ant$ecotroph_gfdl_historical[[1]],
    brick_data_annual_subset_mean_ant$ecotroph_ipsl_historical[[1]],
    brick_data_annual_subset_mean_ant$macroecological_gfdl_historical[[1]],
    brick_data_annual_subset_mean_ant$macroecological_ipsl_historical[[1]],
    brick_data_annual_subset_mean_ant$zoomss_gfdl_historical[[1]],
    brick_data_annual_subset_mean_ant$zoomss_ipsl_historical[[1]])

mean_1<-stackApply(trial_1, indices =  rep(1,nlayers(trial_1)), fun = "mean", na.rm = T)

# ### checking means and whether they make sense given single models - phase 1
# trial_1 # for WS_5_60: 
# (7.246053+53.081787+70.417564+39.263441+306.929849+112.348098+61.729895+53.103343)/8 # 88 does not seem to make sense given max value in map is ~120. But this is becasue for some grid cell only a subset of models is available (hence e.g. (70+330+112)/3 = 123). these are mostly GFDL models. This means that na.rm=F would remove these grid cells. in terms of patterns that does not make much difference but in terms of max values it does and a this is due to a few coastal cells. See Phase 2 below.     
#
# trial_1 ## for 5_60 (or vice versa) 
# (7+50+300+35+250+100+60+50)/8 # model n.3 is the only different one. 

# do size class 6 where boats and macroecological are not available

trial_6<-stack(
    brick_data_annual_subset_mean_ant$apecosm_ipsl_historical[[6]],
    brick_data_annual_subset_mean_ant$dbpm_ipsl_historical[[6]], 
    brick_data_annual_subset_mean_ant$ecotroph_gfdl_historical[[6]],
    brick_data_annual_subset_mean_ant$ecotroph_ipsl_historical[[6]],
    brick_data_annual_subset_mean_ant$zoomss_gfdl_historical[[6]], 
    brick_data_annual_subset_mean_ant$zoomss_ipsl_historical[[6]])

mean_6<-stackApply(trial_6, indices =  rep(1,nlayers(trial_6)), fun = "mean", na.rm = T)

# append all data together 
mean_all<-append(mean, mean_6)
mean_all<-append(mean_1, mean_all)
mean_all[sapply(mean_all, is.null)] <- NULL

# CHECK - OK
# calculate mean using other approach for size class 2 and compare results to above approach

trial<-lapply(brick_data_annual_subset_mean_ant, "[[", 2)
trial[sapply(trial, is.null)] <- NULL

trial2<-lapply(trial, function(x) as.data.frame(rasterToPoints(x))) 
trial2$model = gsub("[^a-zA-Z]", "", rownames(trial2))

trial3<-do.call(rbind.data.frame, trial2)
trial4<-trial3 %>%
  group_by(x,y) %>%
  summarise(mean = mean(index_1, na.rm=FALSE)) %>%
  ungroup()

trial_raster<-rasterFromXYZ(trial4)

# same outputs but trial_raster includes higher lats (empty...)
plot(trial_raster)
plot(mean_all[[2]])

# ### checking means and whether they make sense given single models - phase 2
# filter(trial4, trial4$mean == max(trial4$mean))
# #      x     y  mean
# #   <dbl> <dbl> <dbl>
# # 1  172. -77.5  134.
# 
# filter(trial3, x == 172.5, y == -77.5)
# #                                          x     y   index_1
# # ecotroph_gfdl_historical.4208        172.5 -77.5  51.87132
# # macroecological_gfdl_historical.4211 172.5 -77.5 304.14745
# # zoomss_gfdl_historical.4211          172.5 -77.5  45.57561
# 
# # However, here na.rm=TRUE is == to na.rm=FALSE because, by default, rasterToPoints() omits the NA values. options to change this but not straightforward...  

##### check outputs and reason in comparison with size-spectrum paper 
# cannot compare to these ones as they are up to -40 lat
plot(mean_all[[1]])    
plot(mean_all[[2]])   
plot(mean_all[[3]])   
plot(mean_all[[4]])    
plot(mean_all[[5]])    
plot(mean_all[[6]])  

```

## save means 

```{r}

save(mean_all, file = "/home/ubuntu/Extract_SouthernOcean/Tmp_data/01_extract_modelled_data_SO_means.RData")

# 01_extract_modelled_data_SO_WS_5_60.RData = original sent to Philip as first draft # differences using the latitudinal bands approach are explained inside the function and are due to EcoTroph but nothing to consider further. The final data is OK

```
