---
title: "Extract data from Earth System modelas for the Southern Ocean"
author: "Camilla Novaglio"
date: "01/05/2023"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### AIM

Extract mean values of phytoplankton, and zooplankton the period 1990-1999 across ESM used in FishMIP for comparison with yang et al. 2022. 

phyc-vint: Integrated Total Phytoplankton Carbon Concentration 
zooc-vint: Integrated Total Zooplankton Carbon Concentration 

**Alternative option** for better comparison: 

zmeso-vint: Integrated Mesozooplankton/Large Zoo Carbon Concentration 

### Data source 

Dowloaded on 30/04/2023 - Phyto and zooplnkton are downloaded from DKRZ as per ISIMIP3b protocl: phytoc and zooc (total concentration - https://fish-mip.github.io/protocols.html), vertically integrated. 

scp -r b381217@levante.dkrz.de:/work/bb0820/ISIMIP/ISIMIP3b/InputData/climate/ocean/uncorrected/global/monthly/historical/GFDL-ESM4/gfdl-esm4_r1i1p1f1_historical_phyc-vint_60arcmin_global_monthly_1850_2014.nc /rd/gem/private/users/camillan/Extract_SouthernOcean_Data

scp -r b381217@levante.dkrz.de:/work/bb0820/ISIMIP/ISIMIP3b/InputData/climate/ocean/uncorrected/global/monthly/historical/GFDL-ESM4/gfdl-esm4_r1i1p1f1_historical_zooc-vint_60arcmin_global_monthly_1850_2014.nc /rd/gem/private/users/camillan/Extract_SouthernOcean_Data

scp -r b381217@levante.dkrz.de:/work/bb0820/ISIMIP/ISIMIP3b/InputData/climate/ocean/uncorrected/global/monthly/historical/IPSL-CM6A-LR/ipsl-cm6a-lr_r1i1p1f1_historical_phyc-vint_60arcmin_global_monthly_1850_2014.nc /rd/gem/private/users/camillan/Extract_SouthernOcean_Data

scp -r b381217@levante.dkrz.de:/work/bb0820/ISIMIP/ISIMIP3b/InputData/climate/ocean/uncorrected/global/monthly/historical/IPSL-CM6A-LR/ipsl-cm6a-lr_r1i1p1f1_historical_zooc-vint_60arcmin_global_monthly_1850_2014.nc /rd/gem/private/users/camillan/Extract_SouthernOcean_Data

### *TO CHECK* 

**Check better** file orientation (upside-down?) and check same things as for 01

Strange time dimention now adjusted but this is my interpretation. **Ask Matthias**

Final values are mean mol m-2 for the period 1990-1999 -> for models providing monthly values these are 1) annual means and 2) period means; for models providing annual values these are period means. These outputs have no time dimension (i.e. they represent a stock and not a rate, as opposed to NPP which is in s-1 or to catches which should be in e.g. year-1). Hence, here to get annual values, we average instead of sum across months. For NPP and catch we would sum across months to get yearly valuee as these are rates with a time dimension.      

To transform from mol m-2 to g C m-2 I did X12.01

For better comparison use **zmeso-vint**? I don't totally understand this output -e.g. which size included and how this has been created - and I would use total zooplankton instead.  

Values not decreasing from phyto-zoo-tcb? Should we do **ESMs+MEMs driven by phyto/zoo only and separately**? 

### Set environment 

```{r environment, echo = FALSE}

rm(list=ls())

library(ncdf4) 
library(raster) 
library(ggplot2) 
library(data.table)
library(dtplyr)
library(ncdf4.helpers)
library(tictoc)
library(dplyr, warn.conflicts = FALSE)
library(broom)
library(tibble)
library(tidyr)
library(purrr)
library(stringr)
library(parallel)
library(pbapply)

select<-dplyr::select
summarise<-dplyr::summarise

# load functions 
source("/home/ubuntu/Extract_SouthernOcean/R/Helper_functions_SO.r")

```

### Read in files and extract modelled data at global scale

```{r step 1 to 2}

# create file names ----
dir<-"/rd/gem/private/users/camillan/Extract_SouthernOcean_Data" 

# define them manually as they are only 4
netcdf_phyto= list("gfdl-esm4_r1i1p1f1_historical_phyc-vint_60arcmin_global_monthly_1850_2014.nc",
                   "ipsl-cm6a-lr_r1i1p1f1_historical_phyc-vint_60arcmin_global_monthly_1850_2014.nc")

netcdf_zoo= list("gfdl-esm4_r1i1p1f1_historical_zooc-vint_60arcmin_global_monthly_1850_2014.nc",
                 "ipsl-cm6a-lr_r1i1p1f1_historical_zooc-vint_60arcmin_global_monthly_1850_2014.nc")

# apply function in // ----

phyto_ant<-mclapply(netcdf_phyto, function(x) extract_antarctica_inputs(x), mc.cores = detectCores()-2)
# are names OK?? TO CHECK 
names(phyto_ant) <- netcdf_phyto

zoo_ant<-mclapply(netcdf_zoo, function(x) extract_antarctica_inputs(x), mc.cores = detectCores()-2)
names(zoo_ant) <- netcdf_zoo

```

### save data

```{r}

save(phyto_ant, zoo_ant, file = "/home/ubuntu/Extract_SouthernOcean/Tmp_data/02_extract_input_SO.RData")

```

### multimodel means 

```{r}

rm(list=ls())
load("/home/ubuntu/Extract_SouthernOcean/Tmp_data/02_extract_input_SO.RData")

# combine outputs and average
file_phyto<-stack(phyto_ant[[1]], phyto_ant[[2]])
mean_phyto<-stackApply(file_phyto, indices =  rep(1,nlayers(file_phyto)), fun = "mean", na.rm = F)

plot(mean_phyto) 

file_zoo<-stack(zoo_ant[[1]], zoo_ant[[2]])
mean_zoo<-stackApply(file_zoo, indices =  rep(1,nlayers(file_zoo)), fun = "mean", na.rm = F)

# CHECK 
# calculate mean using other approach for size class 2 and compare results to above approach 

trial<-list(zoo_ant$`gfdl-esm4_r1i1p1f1_historical_zooc-vint_60arcmin_global_monthly_1850_2014.nc`, zoo_ant$`ipsl-cm6a-lr_r1i1p1f1_historical_zooc-vint_60arcmin_global_monthly_1850_2014.nc`)

trial2<-lapply(trial, function(x) as.data.frame(rasterToPoints(x)))
for(i in 1:length(trial2)){trial2[[i]]$model = names(trial2)[i]}

trial3<-do.call(rbind.data.frame, trial2)
trial4<-trial3 %>% 
  group_by(x,y) %>% 
  summarise(mean = mean(index_1, na.rm=TRUE)) %>% 
  ungroup

trial_raster<-rasterFromXYZ(trial4)

# same outputs but trial_raster includes higher lats (empty...)
plot(trial_raster)
plot(mean_zoo)

```

## save means 

```{r}

save(mean_phyto, mean_zoo, file = "/home/ubuntu/Extract_SouthernOcean/Tmp_data/02_extract_input_SO_means_narmF.RData")

```
