---
title: "Extract data from Earth System modelas for the Southern Ocean"
author: "Camilla Novaglio"
date: "01/05/2023"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### AIM

Extract mean values of phytoplankton, and zooplankton the period 1990-1999 across ESM used in FishMIP for comparison with yang et al. 2022. 

phyc-vint: Integrated Total Phytoplankton Carbon Concentration 
zooc-vint: Integrated Total Zooplankton Carbon Concentration 

### Data source 

Dowloaded on 30/04/2023 - Phyto and zooplnkton are downloaded from DKRZ as per ISIMIP3b protocol: phyc and zooc (total concentration - https://fish-mip.github.io/protocols.html), vertically integrated. 

We consider ISIMIP3b data as we have MEM outputs linked to these inputs for ISIMIP3b.

scp -r b381217@levante.dkrz.de:/work/bb0820/ISIMIP/ISIMIP3b/InputData/climate/ocean/uncorrected/global/monthly/historical/GFDL-ESM4/gfdl-esm4_r1i1p1f1_historical_phyc-vint_60arcmin_global_monthly_1850_2014.nc /rd/gem/private/users/camillan/Extract_SouthernOcean_Data

scp -r b381217@levante.dkrz.de:/work/bb0820/ISIMIP/ISIMIP3b/InputData/climate/ocean/uncorrected/global/monthly/historical/GFDL-ESM4/gfdl-esm4_r1i1p1f1_historical_zooc-vint_60arcmin_global_monthly_1850_2014.nc /rd/gem/private/users/camillan/Extract_SouthernOcean_Data

scp -r b381217@levante.dkrz.de:/work/bb0820/ISIMIP/ISIMIP3b/InputData/climate/ocean/uncorrected/global/monthly/historical/IPSL-CM6A-LR/ipsl-cm6a-lr_r1i1p1f1_historical_phyc-vint_60arcmin_global_monthly_1850_2014.nc /rd/gem/private/users/camillan/Extract_SouthernOcean_Data

scp -r b381217@levante.dkrz.de:/work/bb0820/ISIMIP/ISIMIP3b/InputData/climate/ocean/uncorrected/global/monthly/historical/IPSL-CM6A-LR/ipsl-cm6a-lr_r1i1p1f1_historical_zooc-vint_60arcmin_global_monthly_1850_2014.nc /rd/gem/private/users/camillan/Extract_SouthernOcean_Data

Downloaded again on 25/08/2023 to check any change to the time dimension

### download mesozooplankton 

scp -r b381217@levante.dkrz.de:/work/bb0820/ISIMIP/ISIMIP3b/InputData/climate/ocean/uncorrected/global/monthly/historical/GFDL-ESM4/gfdl-esm4_r1i1p1f1_historical_zmeso-vint_60arcmin_global_monthly_1850_2014.nc /rd/gem/private/users/camillan/Extract_SouthernOcean_Data

scp -r b381217@levante.dkrz.de:/work/bb0820/ISIMIP/ISIMIP3b/InputData/climate/ocean/uncorrected/global/monthly/historical/IPSL-CM6A-LR/ipsl-cm6a-lr_r1i1p1f1_historical_zmeso-vint_60arcmin_global_monthly_1850_2014.nc /rd/gem/private/users/camillan/Extract_SouthernOcean_Data

### Checks 

Time dimension now adjusted. Asked Matthias, interpretation is OK - starting value is Jan 1850.  

Final values are mean mol m-2 for the period indicated below -> for models providing monthly values these are 1) annual means and 2) period means; for models providing annual values these are period means. These outputs have no time dimension (i.e. they represent a stock and not a rate, as opposed to NPP which is in s-1 or to catches which should be in e.g. year-1). Hence, here to get annual values, we average instead of sum across months. For NPP and catch we would sum across months to get yearly valuee as these are rates with a time dimension.      

To transform from mol m-2 to g C m-2 I did X12

For better comparison use zmeso-vint: Integrated Mesozooplankton/Large Zoo Carbon Concentration? Discussed with Ryan - not fully understanding this output -e.g. which size included and how this has been created - would use total zooplankton instead as the one commonly used in FishMIP papers and plots and specify in figs. checked mesozooplankton too: different patterns and values as expected,  best to use zooplankton as we are more interested in transfer of biomass across trophic levels. 

### Set environment 

```{r environment, echo = FALSE}

rm(list=ls())

library(ncdf4) 
library(raster) 
library(ggplot2) 
library(data.table)
library(dtplyr)
library(ncdf4.helpers)
library(tictoc)
library(dplyr, warn.conflicts = FALSE)
library(broom)
library(tibble)
library(tidyr)
library(purrr)
library(stringr)
library(parallel)
library(pbapply)

select<-dplyr::select
summarise<-dplyr::summarise

# load functions 
source("/home/ubuntu/Extract_SouthernOcean/R/Helper_functions_SO.r")

# plankton options (latest agreed version is summer and zoo and Hyear 1995)
season = "year" # "year"# "summer"
zoopl = "zoo" # "zoo" # "meso"

```

### Read in files and extract modelled data at global scale

```{r step 1 to 2}

# create file names ----
dir<-"/rd/gem/private/users/camillan/Extract_SouthernOcean_Data" 

# define them manually as they are only 4
netcdf_phyto= c("gfdl-esm4_r1i1p1f1_historical_phyc-vint_60arcmin_global_monthly_1850_2014.nc",
                "ipsl-cm6a-lr_r1i1p1f1_historical_phyc-vint_60arcmin_global_monthly_1850_2014.nc")

if(zoopl == "zoo"){
  netcdf_zoo= c("gfdl-esm4_r1i1p1f1_historical_zooc-vint_60arcmin_global_monthly_1850_2014.nc",
                "ipsl-cm6a-lr_r1i1p1f1_historical_zooc-vint_60arcmin_global_monthly_1850_2014.nc")
}else if (zoopl == "meso"){
  netcdf_zoo= c("gfdl-esm4_r1i1p1f1_historical_zmeso-vint_60arcmin_global_monthly_1850_2014.nc",
                "ipsl-cm6a-lr_r1i1p1f1_historical_zmeso-vint_60arcmin_global_monthly_1850_2014.nc")
}

# apply function in // ----

history_year1 = 1995 # 1998 (20-years closed to the empirical data) # for empirical: December to March 1998-2020 for phyto and unknown for meso. # 1990
history_year2 = 2014 # 2014 (last year of historical projections) #1999
future_year1 = NA
future_year2 = NA

phyto_ant<-mclapply(netcdf_phyto, function(x) extract_antarctica_inputs(x, history_year1 =history_year1, history_year2=history_year2,future_year1 = future_year1,future_year2=future_year2, season = season), mc.cores = detectCores()-2)

names(phyto_ant) <- netcdf_phyto

zoo_ant<-mclapply(netcdf_zoo, function(x) extract_antarctica_inputs(x, history_year1 =history_year1, history_year2=history_year2,future_year1 = future_year1,future_year2=future_year2, season = season), mc.cores = detectCores()-2)

names(zoo_ant) <- netcdf_zoo

```

### save data 

```{r}

## NOTE - if you run this you need to adjust as per naming below 
# if(season == "summer"){
#   save(phyto_ant, zoo_ant, file = paste("/home/ubuntu/Extract_SouthernOcean/Tmp_data/", Sys.Date(), "_02_extract_input_SO_summer.RData", sep =""))
# }else if(season == "year") {
#   save(phyto_ant, zoo_ant, file = paste("/home/ubuntu/Extract_SouthernOcean/Tmp_data/", Sys.Date(), "_02_extract_input_SO.RData", sep = ""))
# }

```

### multimodel means 

```{r}

# rm(list=ls())
# 
# if(season == "summer"){
#   load("/home/ubuntu/Extract_SouthernOcean/Tmp_data/02_extract_input_SO_summer.RData")
# }else if(season == "year") {
#   load("/home/ubuntu/Extract_SouthernOcean/Tmp_data/02_extract_input_SO.RData")
# }

removeNA = "narmF" # narmT or narmF

file_phyto<-stack(phyto_ant[[1]], phyto_ant[[2]])
file_zoo<-stack(zoo_ant[[1]], zoo_ant[[2]])

if(removeNA == "narmT"){

  # combine outputs and average
  mean_phyto<-stackApply(file_phyto, indices =  rep(1,nlayers(file_phyto)), fun = mean, na.rm = TRUE)
  mean_zoo<-stackApply(file_zoo, indices =  rep(1,nlayers(file_zoo)), fun = mean, na.rm = TRUE)

  # # CHECK
  # # calculate mean using other approach for size class 2 and compare results to above approach
  # 
  # trial<-list(zoo_ant$`gfdl-esm4_r1i1p1f1_historical_zooc-vint_60arcmin_global_monthly_1850_2014.nc`, zoo_ant$`ipsl-cm6a-lr_r1i1p1f1_historical_zooc-vint_60arcmin_global_monthly_1850_2014.nc`)
  # 
  # trial2<-lapply(trial, function(x) as.data.frame(rasterToPoints(x)))
  # for(i in 1:length(trial2)){trial2[[i]]$model = names(trial2)[i]}
  # 
  # trial3<-do.call(rbind.data.frame, trial2) %>%
  #   group_by(x,y) %>%
  #   summarise(mean = mean(index_1, na.rm=TRUE)) %>%
  #   ungroup
  # 
  # trial_raster<-rasterFromXYZ(trial3)
  # 
  # # same outputs 
  # plot(trial_raster)
  # plot(mean_zoo)

}else if (removeNA == "narmF"){
  
  mean_phyto<-stackApply(file_phyto, indices =  rep(1,nlayers(file_phyto)), fun = mean, na.rm = FALSE)
  mean_zoo<-stackApply(file_zoo, indices =  rep(1,nlayers(file_zoo)), fun = mean, na.rm = FALSE)

}

```

## save means 

```{r}

save(mean_phyto, mean_zoo, file = paste("/home/ubuntu/Extract_SouthernOcean/Tmp_data/",  Sys.Date(), "_02_extract_input_SO_means_", removeNA, "_", season, "_zooType_", zoopl, "_Hyear_",history_year1, ".RData", sep = ""))

# old version - yearly, 1990s, zoo outputs are saved with these names

# if(season == "year"){
#   if(removeNA == "narmT"){
#     save(mean_phyto, mean_zoo, file = pate( "/home/ubuntu/Extract_SouthernOcean/Tmp_data/",  Sys.Date(), "_02_extract_input_SO_means.RData", sep = ""))
#   }else if(removeNA == "narmF"){
#     save(mean_phyto, mean_zoo, file = paste("/home/ubuntu/Extract_SouthernOcean/Tmp_data/", Sys.Date(),"_02_extract_input_SO_means_narmF.RData", sep = ""))
#   }
# }else if (season == "summer")  {
#     if(removeNA == "narmT"){
#     save(mean_phyto, mean_zoo, file = paste( "/home/ubuntu/Extract_SouthernOcean/Tmp_data/",  Sys.Date(),"_02_extract_input_SO_means_summer.RData", sep = ""))
#   }else if(removeNA == "narmF"){
#     save(mean_phyto, mean_zoo, file = paste( "/home/ubuntu/Extract_SouthernOcean/Tmp_data/", Sys.Date(),"_02_extract_input_SO_means_narmF_summer.RData", sep = ""))
#   }
# }

```
