---
title: "PLot maps of MEMs and ESMs for the Southern Ocean"
author: "Camilla Novaglio"
date: "01/05/2023"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### NOTEs from editor 

in all the figures, we will need to avoid the use of the rainbow color scale. This is another benefit of re-plotting. Instead, please use a single colour intensity scale or otherwise see https://www.nature.com/articles/s41467-020-19160-7 for options of other scientific colour scales.

In general, something to be aware of: all figures must be provided in high resolution (min 300dpi), editable format (PDF, EPS or AI). This is so that the art editors can modify them to conform with style, later in the process. If figures are produced one program and modified in another before being coverted to EPS/PDF, the ‘placed’ images are not editable. PNG or PJEG are not editable. I mention this now to save time later as you may well revise the figures based on referee comments.

### Set environment 

```{r environment, echo = FALSE}

rm(list=ls())

library(raster) 
library(ggplot2) 
library(dplyr)
library(sf)
library(rnaturalearth)

select<-dplyr::select
summarise<-dplyr::summarise

my_theme_map<-theme_bw()+ 
  theme(
    text = element_text(size = 9), 
    plot.title = element_text(size = 9),
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    legend.title=element_text(size = 8), 
    legend.text=element_text(size = 7),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), 
    panel.background = element_blank(),
    plot.background = element_rect(fill="white"),
    panel.border = element_blank(),
    legend.key.height = unit(0.2, "cm"), # this becomes width if you choose bottom position
    legend.key.width = unit(0.5, "cm"), 
    legend.position= 'bottom' # c(0.5,0.6) # in middle
)


# load functions 
source("/home/ubuntu/Extract_SouthernOcean/R/Plotting_functions_SO.r")

```

## Load data means 

```{r}

load("/home/ubuntu/Extract_SouthernOcean/Tmp_data/01_extract_modelled_data_SO_means_narmF.RData")
# ls()
# mean_all

load("/home/ubuntu/Extract_SouthernOcean/Tmp_data/02_extract_input_SO_means_narmF.RData")
# ls()
# mean_phyto
# mean_zoo

load("/home/ubuntu/Extract_SouthernOcean/Tmp_data/03_work_with_empirical.RData")
# phyt_raster ### WARNING - change names to better distinguish modelled vs empirical
# zoop_raster

```

## Map means 

```{r}

# extract world map and prepare in circumpolar projections:  

#### Southern Ocean 
# antmap <- ne_download(scale = 110, type = 'countries') # dowloaded as spatialPoligonDataframe
# # # if you want to plot in lat/lon 
# # antmap_df <- fortify(antmap)
# pr <- "+proj=laea +lat_0=-90 +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0"   
# antmap <- spTransform(antmap, CRS(pr)) 
# antmap_df <- fortify(antmap)

#### Crop region you are interested in (https://www.r-bloggers.com/2019/04/zooming-in-on-maps-with-sf-and-ggplot2/)
antmap <- ne_countries(scale = 'medium', type = 'map_units',returnclass = 'sf') # dowloaded as simple feature
antmap2 <- st_crop(antmap, xmin = -180, xmax = 180, ymin = -90, ymax = -40) # if you do -35 you get the same land plot as Yang. et al. 2022
# WARNING - not understanding why the problem is solved here.... 
# # problem here - solved using https://github.com/r-spatial/sf/issues/1902
sf_use_s2(FALSE)
antmap2 <- st_make_valid(antmap)
antmap2 <- st_crop(antmap, xmin = -180, xmax = 180, ymin = -90, ymax = -50)
ggplot() + geom_sf(data = antmap2) + theme_bw()

# from sf to spatialPoligonDataframe 
antmap <- as_Spatial(antmap2)
pr <- "+proj=laea +lat_0=-90 +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0"   
antmap <- spTransform(antmap, CRS(pr)) 
antmap_df <- fortify(antmap)

# convert rasterLayer to sf object and consider circumpolar projections within function
circumpolarCRS <- CRS("+proj=laea +lat_0=-90 +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0")

###### produce maps 

## WARNING - might be better to loop to avoid mistakes ..... 
## TO FIX IN FUNCTION aes(fill = index_1) for all modelled data and aes(fill = layer) fo empirical data - I don't know why and how to change this in previous code ...

####### Empirical data - to add to figs below 





source("/home/ubuntu/Extract_SouthernOcean/R/Plotting_functions_SO.r")
# phytoplankton  
min = 0
max = 8.75 # as per Yang et al. 2022
legend_name = "g C m-2"

phyto_emp_plot<-plot_SO(wmap_df = antmap_df, data_to_plot = phyt_raster, proj = circumpolarCRS, legend_name = legend_name, min = min, max = max)
phyto_emp_plot

# phyto_emp_plot<-phyto_emp_plot+ggtitle("A) Phytoplankton empirical")

source("/home/ubuntu/Extract_SouthernOcean/R/Plotting_functions_SO.r")
# zooplankton  
min = 0
max = 25 # as per Yang et al. 2022
legend_name = "g C m-2"

zoo_emp_plot<-plot_SO(wmap_df = antmap_df, data_to_plot = zoop_raster, proj = circumpolarCRS, legend_name = legend_name, min = min, max = max)
zoo_emp_plot

# zoo_emp_plot<-zoo_emp_plot+ggtitle("B) Zooplankton empirical")











####### FIG 1 

# phytoplankton  
# min = 0
# max = 8.75 # as per Yang et al. 2022 
legend_name = "g C m-2"

phyto_plot<-plot_SO(wmap_df = antmap_df, data_to_plot = mean_phyto, proj = circumpolarCRS, legend_name = legend_name, min = mean_phyto@data@min, max = mean_phyto@data@max)

phyto_plot<-phyto_plot+ggtitle("A) Phytoplankton")

# # check map
# phyto_plot
# plot(mean_phyto)

# mesozooplankton limits 
# min = 0
# max = 25 # as per Yang et al. 2022 
legend_name = "g C m-2"
  
plot(mean_zoo)

zoo_plot<-plot_SO(wmap_df = antmap_df, data_to_plot = mean_zoo, proj = circumpolarCRS, legend_name = legend_name, min = mean_zoo@data@min, max = mean_zoo@data@max) # almost empty high number so adjustng the range

zoo_plot<-zoo_plot+ggtitle("B) Zooplankton")

###### FIG S1 

min = 0.5
max = 5

# phytoplankton  
legend_name = "g C m-2"

phyto_plot2<-plot_SO(wmap_df = antmap_df, data_to_plot = mean_phyto, proj = circumpolarCRS, legend_name = legend_name, min = min, max = max)

phyto_plot2<-phyto_plot2+ggtitle("A) Phytoplankton")

# Zooplankton 
legend_name = "g C m-2"
  
zoo_plot2<-plot_SO(wmap_df = antmap_df, data_to_plot = mean_zoo, proj = circumpolarCRS, legend_name = legend_name, min = min, max = max) 

zoo_plot2<-zoo_plot2+ggtitle("B) Zooplankton")

###### FIG S2 

# tcb limits size equal to all size classes 
legend_name = "g C m-2"

mean_all[[3]]@data@min
mean_all[[3]]@data@max
plot(mean_all[[1]])

size1_plot<-plot_SO(wmap_df = antmap_df, data_to_plot = mean_all[[1]], proj = circumpolarCRS, legend_name = legend_name, min = min, max = max)

size1_plot<-size1_plot+ggtitle("A) Total Consumers Biomass 1-10g")

plot(mean_all[[2]])

size2_plot<-plot_SO(wmap_df = antmap_df, data_to_plot = mean_all[[2]], proj = circumpolarCRS, legend_name = legend_name, min = min, max = max)

size2_plot<-size2_plot+ggtitle("B) Total Consumers Biomass 10-100g")

plot(mean_all[[3]])

size3_plot<-plot_SO(wmap_df = antmap_df, data_to_plot = mean_all[[3]], proj = circumpolarCRS, legend_name = legend_name, min = min, max = max)

size3_plot<-size3_plot+ggtitle("C) Total Consumers Biomass 100g-1kg")

plot(mean_all[[4]])

size4_plot<-plot_SO(wmap_df = antmap_df, data_to_plot = mean_all[[4]], proj = circumpolarCRS, legend_name = legend_name, min = min, max = max)

size4_plot<-size4_plot+ggtitle("D) Total Consumers Biomass 1-10kg")

plot(mean_all[[5]])

size5_plot<-plot_SO(wmap_df = antmap_df, data_to_plot = mean_all[[5]], proj = circumpolarCRS, legend_name = legend_name, min = min, max = max)

size5_plot<-size5_plot+ggtitle("E) Total Consumers Biomass 10-100kg")

size6_plot<-plot_SO(wmap_df = antmap_df, data_to_plot = mean_all[[6]], proj = circumpolarCRS, legend_name = legend_name, min = min, max = max)

size6_plot<-size6_plot+ggtitle("F) Total Consumers Biomass 100kg-1t")

```

### Save fianl plots 

```{r}

library(patchwork)

# WARNINIG 
# try na.rm = F when mean across models to explore coastal patterns where only GFDL-driven models are available. the yellow coastal cells disapper. 
# phyto and zoo (Fig.5) are different when doing na.rm = T or F because the scale considered changes (i.e. smallest and largest values are removed when na.rm = F for zoo in particular). range of values are smaller and closer together.  

# phyto and zooplankton fig 5
jpeg("/home/ubuntu/Extract_SouthernOcean/Output/Fig5_narmF.jpg", width = 5, height = 5, units = "in", res = 300, bg = "transparent")
phyto_plot+zoo_plot
dev.off()

# phyto and zooplankton fig S1 
p_final<-phyto_plot2+zoo_plot2+
  plot_layout(ncol = 2, nrow = 1, guides = 'collect')

jpeg("/home/ubuntu/Extract_SouthernOcean/Output/FigS1_narmF.jpg", width = 10, height = 10, units = "in", res = 300, bg = "transparent")
p_final
dev.off()

# tcb by size classes FigS2
p_final<-size1_plot+size2_plot+
  size3_plot+size4_plot+
  size5_plot+size6_plot+
  plot_layout(ncol = 3, nrow = 2, guides = 'collect')

jpeg("/home/ubuntu/Extract_SouthernOcean/Output/FigS2_narmF.jpg", width = 10, height = 10, units = "in", res = 300, bg = "transparent")
p_final
dev.off()

```


