---
title: "PLot maps of MEMs and ESMs for the Southern Ocean"
author: "Camilla Novaglio"
date: "01/05/2023"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### NOTEs from editor 

Avoid the use of the rainbow color scale. Instead, please use a single colour intensity scale or otherwise see https://www.nature.com/articles/s41467-020-19160-7 for options of other scientific colour scales.

viridis is accepted for maps and colorbrewer too. 

print pdf - units are inches by def (7) and pdf is vector graphic so no need for resolution to be defined. 

## Set environment 

```{r environment, echo = FALSE}

rm(list=ls())

library(raster) 
library(ggplot2) 
library(dplyr)
library(sf)
library(rnaturalearth)
library(patchwork)
library(colorspace)

select<-dplyr::select
summarise<-dplyr::summarise

my_theme_map<-theme_bw()+ 
  theme(
    text = element_text(size = 9), 
    plot.title = element_text(size = 9),
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    legend.title=element_text(size = 8), 
    legend.text=element_text(size = 7),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), 
    panel.background = element_blank(),
    plot.background = element_rect(fill="white"),
    panel.border = element_blank(),
    legend.key.height = unit(0.2, "cm"), # this becomes width if you choose bottom position
    legend.key.width = unit(1, "cm"),
    legend.position= 'bottom' # c(0.5,0.6) # in middle
)

my_theme_map2<-theme_bw()+ 
  theme(
    text = element_text(size = 9), 
    plot.title = element_text(size = 9),
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    legend.title=element_text(size = 8), 
    legend.text=element_text(size = 7),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), 
    panel.background = element_blank(),
    plot.background = element_rect(fill="white"),
    panel.border = element_blank(),
    legend.key.height = unit(0.7, "cm"), # this becomes width if you choose bottom position
    legend.key.width = unit(0.2, "cm")
  )    


# load functions 
source("/home/ubuntu/Extract_SouthernOcean/R/Plotting_functions_SO.r")

```

## set arguments 

```{r}

# loading data 
removeNA = "narmF" # options narmT and narmF

# trophic amplification 
ltl_type = "sum" # options sum, phyto and zoop. Philip: sum preferred. 

# mapping color 
color_scale = "viridis" # options viridis and monochromatic and divergent (used for throphic amplification)

# mapping number of breaks
breaks_vector_n = 7

# mapping scale 
scale_Fig0 = "raw" # not needed
scale_Fig4 = "log" # new plot for Philip - with log scale and same legend
scale_FigS1 = "raw" # basically same as Fig 5 but with raw values - should you have different legends here?
scale_FigS2 = "log" # new plot for Philip - with log scale to go with Fig4 and same legend
scale_FigS3 = "raw" # new plot for Philip - scale needs to be raw - and same legend? 

# mapping limits
min_Fig0 = 0
max_Fig0_phyt = 8.7
max_Fig0_zoop = 25

min_Fig4 = 0.5 # if up to each dataset limits: e.g.  min = phyt_raster@data@min
max_Fig4 = 5 # max = min = phyt_raster@data@max

min_FigS1 = 0.5
max_FigS1 = 5

min_FigS2 = 0.5
max_FigS2 = 5

min_FigS3 = 0.5 # 0.5 means that the higher trophic level is 1/2 the lower trophic level.
max_FigS3 = 1.5 # 2 means that the higher trophic level is double the lower trophic level. 

```

## Load data means 

```{r}

# phyto and zoo are different when doing na.rm = T or F because the scale considered changes (i.e. smallest and largest values are removed when na.rm = F for zoo in particular). range of values are smaller and closer together.

if(removeNA == "narmT"){
  load("/home/ubuntu/Extract_SouthernOcean/Tmp_data/01_extract_modelled_data_SO_means.RData")
  load("/home/ubuntu/Extract_SouthernOcean/Tmp_data/02_extract_input_SO_means.RData")
  load("/home/ubuntu/Extract_SouthernOcean/Tmp_data/03_work_with_empirical.RData")
}else if (removeNA == "narmF"){
  load("/home/ubuntu/Extract_SouthernOcean/Tmp_data/01_extract_modelled_data_SO_means_narmF.RData")
  load("/home/ubuntu/Extract_SouthernOcean/Tmp_data/02_extract_input_SO_means_narmF.RData")
  load("/home/ubuntu/Extract_SouthernOcean/Tmp_data/03_work_with_empirical.RData")
}

```

## calculate trophic amplification

Philip: Could you plot the amplification effect for the model. So for each pixel upper trophic level biomass / lower trophic level biomass

```{r}

if(ltl_type == "phyto"){
 ltl<-mean_phyto 
}else if(ltl_type == "zoop"){
  ltl<-mean_zoo
}else if(ltl_type == "sum"){
  ltl<-mean_phyto+mean_zoo
}

# # CHECK
# plot(mean_phyto)
# plot(mean_zoo)
# plot(ltl)
# 
# # CHECK 
# trial1<-as.data.frame(rasterToPoints(mean_phyto))
# trial2<-as.data.frame(rasterToPoints(mean_zoo))
# trial3<-as.data.frame(rasterToPoints(ltl))
# head(trial1)
# head(trial2)
# head(trial3)

tl2<-mean_all[[1]]/ltl
tl3<-mean_all[[2]]/mean_all[[1]]
tl4<-mean_all[[3]]/mean_all[[2]]
tl5<-mean_all[[4]]/mean_all[[3]]
tl6<-mean_all[[5]]/mean_all[[4]]
tl7<-mean_all[[6]]/mean_all[[5]]

# # CHECK 
# plot(tl2)
# plot(mean_all[[1]])
# plot(ltl)
# 
# # CHECK
# trial1<-as.data.frame(rasterToPoints(mean_all[[1]]))
# trial2<-as.data.frame(rasterToPoints(ltl))
# trial3<-as.data.frame(rasterToPoints(tl2))
# head(trial1)
# head(trial2)
# head(trial3)

```

## Load land map 

```{r}

# extract world map and prepare in circumpolar projections:  

#### Southern Ocean 
# antmap <- ne_download(scale = 110, type = 'countries') # dowloaded as spatialPoligonDataframe
# # # if you want to plot in lat/lon 
# # antmap_df <- fortify(antmap)
# pr <- "+proj=laea +lat_0=-90 +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0"   
# antmap <- spTransform(antmap, CRS(pr)) 
# antmap_df <- fortify(antmap)

#### Crop region you are interested in (https://www.r-bloggers.com/2019/04/zooming-in-on-maps-with-sf-and-ggplot2/)
antmap <- ne_countries(scale = 'medium', type = 'map_units',returnclass = 'sf') # dowloaded as simple feature
antmap2 <- st_crop(antmap, xmin = -180, xmax = 180, ymin = -90, ymax = -40) # if you do -35 you get the same land plot as Yang. et al. 2022
# WARNING - not understanding why the problem is solved here.... 
# # problem here - solved using https://github.com/r-spatial/sf/issues/1902
sf_use_s2(FALSE)
antmap2 <- st_make_valid(antmap)
antmap2 <- st_crop(antmap, xmin = -180, xmax = 180, ymin = -90, ymax = -40) # -40 should cover bothe empirical and modelled data 
# ggplot() + geom_sf(data = antmap2) + theme_bw()

# convert rasterLayer to sf object and consider circumpolar projections within function
circumpolarCRS <- CRS("+proj=laea +lat_0=-90 +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0")

# # from sf to spatialPoligonDataframe 
# antmap <- as_Spatial(antmap2)
# pr <- "+proj=laea +lat_0=-90 +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0"   
# antmap <- spTransform(antmap, CRS(pr)) 
# antmap_df1 <- fortify(antmap)

#### OR 
# from sf to spatialPoligonDataframe 
antmap <- as_Spatial(antmap2)
antmap <- st_as_sf(antmap) # convert dataframe to sf object 
antmap_df <- st_transform(antmap, crs = st_crs(circumpolarCRS)) # convert sf object to circumpolar Projection

```

## Produce maps of data  

```{r}

### Empirical data check ----

# check patterns are as per Yang 

if(scale_Fig0 == "raw"){
  legend_name = expression(g~C~m^{-2})
}else if(scale_Fig0 == "log"){
  legend_name = expression(log~g~C~m^{-2}) 
}

names(phyt_raster)<-"index_1" 
phyt_raster@data@min
phyt_raster@data@max
check_phyto<-plot_SO(wmap_df = antmap_df, data_to_plot = phyt_raster, proj = circumpolarCRS, legend_name = legend_name, min = min_Fig0, max = max_Fig0_phyt, data_type = "phyto", color_scale = "viridis", theme_map = "my_theme_map", scale = scale_Fig0, breaks_vector_n = breaks_vector_n)
check_phyto<-check_phyto+ggtitle("a) Phytoplankton empirical")

names(zoop_raster)<-"index_1" 
zoop_raster@data@min
zoop_raster@data@max # check this better - it also depends on the nrow in 03
check_zoo<-plot_SO(wmap_df = antmap_df, data_to_plot = zoop_raster, proj = circumpolarCRS, legend_name = legend_name, min = min_Fig0, max = max_Fig0_zoop, data_type = "zoo", color_scale = "viridis", theme_map= "my_theme_map", scale = scale_Fig0, breaks_vector_n = breaks_vector_n)
check_zoo<-check_zoo+ggtitle("b) Zooplankton empirical")

### Empirical data for Fig 5 -----

# Fig 5 lower panel: Depth integrated phytoplankton and zooplankton carbon concentration projections averaged across the 1990-1999 decade and the two Earth System Models, IPSL-CM6A-LR and GFDL-ESM4, providing climate forcing data for FishMIP Marine Ecosystem Models, ISIMIP3b simulation round. Data available at https://data.isimip.org/. The range of values in the legend spans the minimum and the maximum value of phytoplankton and of zooplankton, respectively. For a comparison with biomass for higher trophic levels see  S-Fig.1 and 2.
# adding higher panel with same characteristics. 

if(scale_Fig4 == "raw"){
  legend_name = expression(g~C~m^{-2})
}else if(scale_Fig4 == "log"){
  legend_name = expression(log~g~C~m^{-2}) 
}

names(phyt_raster)<-"index_1" 
phyto_emp_plot<-plot_SO(wmap_df = antmap_df, data_to_plot = phyt_raster, proj = circumpolarCRS, legend_name = legend_name, min = min_Fig4, max = max_Fig4, data_type = "modelled", color_scale = color_scale, theme_map= "my_theme_map2", scale = scale_Fig4, breaks_vector_n = breaks_vector_n)
phyto_emp_plot<-phyto_emp_plot+ggtitle("a) Phytoplankton empirical")
phyto_emp_plot

# zoop_raster_trial<-zoop_raster
# zoop_raster_trial[is.na(zoop_raster_trial)]<-Inf

names(zoop_raster)<-"index_1" 
zoop_emp_plot<-plot_SO(wmap_df = antmap_df, data_to_plot = zoop_raster, proj = circumpolarCRS, legend_name = legend_name, min = min_Fig4, max = max_Fig4, data_type = "modelled", color_scale = color_scale, theme_map= "my_theme_map2", scale = scale_Fig4, breaks_vector_n = breaks_vector_n)
zoop_emp_plot<-zoop_emp_plot+ggtitle("b) Zooplankton empirical")
zoop_emp_plot

# names(zoop_raster_trial)<-"index_1" 
# zoop_emp_plot2<-plot_SO(wmap_df = antmap_df, data_to_plot = zoop_raster_trial, proj = circumpolarCRS, legend_name = legend_name, min = min_Fig4, max = max_Fig4, data_type = "modelled", color_scale = color_scale, theme_map= "my_theme_map2", scale = scale_Fig4, breaks_vector_n = breaks_vector_n)
# zoop_emp_plot2<-zoop_emp_plot+ggtitle("b) Zooplankton empirical")
# zoop_emp_plot2

### ESMs data for Fig 5 -----

phyto_plot<-plot_SO(wmap_df = antmap_df, data_to_plot = mean_phyto, proj = circumpolarCRS, legend_name = legend_name, min = min_Fig4, max = max_Fig4, data_type = "modelled", color_scale = color_scale, theme_map= "my_theme_map2", scale = scale_Fig4, breaks_vector_n = breaks_vector_n)
phyto_plot<-phyto_plot+ggtitle("c) Phytoplankton modelled")

zoo_plot<-plot_SO(wmap_df = antmap_df, data_to_plot = mean_zoo, proj = circumpolarCRS, legend_name = legend_name, min = min_Fig4, max = max_Fig4, data_type = "modelled", color_scale = color_scale, theme_map= "my_theme_map2", scale = scale_Fig4, breaks_vector_n = breaks_vector_n) 
zoo_plot<-zoo_plot+ggtitle("d) Zooplankton modelled")

### empirical data for Fig S1 -----

# Fig S1. Depth integrated a) phytoplankton and b) zooplankton carbon concentration projections averaged across the 1990-1999 decade and the two Earth System Models, IPSL-CM6A-LR and GFDL-ESM4, providing climate forcing data for FishMIP Marine Ecosystem Models, ISIMIP3b simulation round. Data available at https://data.isimip.org/. The range of values is kept between 0.5 and 5 g C m-2 for better comparison between plots and with maps of total consumers biomass (S-Fig.2). The code used to produce this figure is available at https://github.com/Fish-MIP/Extract_SouthernOcean.

if(scale_FigS1 == "raw"){
  legend_name = expression(g~C~m^{-2})
}else if(scale_FigS1 == "log"){
  legend_name = expression(log~g~C~m^{-2}) 
}

phyto_plot_emp2<-plot_SO(wmap_df = antmap_df, data_to_plot = phyt_raster, proj = circumpolarCRS, legend_name = legend_name, min = min_FigS1, max = max_FigS1, data_type = "modelled", color_scale = color_scale, theme_map= "my_theme_map2", scale = scale_FigS1, breaks_vector_n = breaks_vector_n)
phyto_plot_emp2<-phyto_plot_emp2+ggtitle("a) Phytoplankton empirical")

zoo_plot_emp2<-plot_SO(wmap_df = antmap_df, data_to_plot = zoop_raster, proj = circumpolarCRS, legend_name = legend_name, min = min_FigS1, max = max_FigS1, data_type = "modelled", color_scale = color_scale, theme_map= "my_theme_map2", scale = scale_FigS1, breaks_vector_n = breaks_vector_n) 
zoo_plot_emp2<-zoo_plot_emp2+ggtitle("b) Zooplankton empirical")

### ESMs data for Fig S1 -----

phyto_plot2<-plot_SO(wmap_df = antmap_df, data_to_plot = mean_phyto, proj = circumpolarCRS, legend_name = legend_name, min = min_FigS1, max = max_FigS1, data_type = "modelled", color_scale = color_scale, theme_map= "my_theme_map2", scale = scale_FigS1, breaks_vector_n = breaks_vector_n)
phyto_plot2<-phyto_plot2+ggtitle("c) Phytoplankton modelled")

zoo_plot2<-plot_SO(wmap_df = antmap_df, data_to_plot = mean_zoo, proj = circumpolarCRS, legend_name = legend_name, min = min_FigS1, max = max_FigS1, data_type = "modelled", color_scale = color_scale, theme_map= "my_theme_map2", scale = scale_FigS1, breaks_vector_n = breaks_vector_n) 
zoo_plot2<-zoo_plot2+ggtitle("d) Zooplankton modelled")


### empirical data for Fig S1B -----

if(scale_FigS1 == "raw"){
  legend_name = expression(g~C~m^{-2})
}else if(scale_FigS1 == "log"){
  legend_name = expression(log~g~C~m^{-2}) 
}

phyto_plot_emp2_S1B<-plot_SO(wmap_df = antmap_df, data_to_plot = phyt_raster, proj = circumpolarCRS, legend_name = legend_name, min = phyt_raster@data@min, max = phyt_raster@data@max, data_type = "modelled", color_scale = color_scale, theme_map= "my_theme_map2", scale = scale_FigS1, breaks_vector_n = breaks_vector_n)
phyto_plot_emp2_S1B<-phyto_plot_emp2_S1B+ggtitle("a) Phytoplankton empirical")

zoo_plot_emp2_S1B<-plot_SO(wmap_df = antmap_df, data_to_plot = zoop_raster, proj = circumpolarCRS, legend_name = legend_name, min = zoop_raster@data@min, max = zoop_raster@data@max, data_type = "modelled", color_scale = color_scale, theme_map= "my_theme_map2", scale = scale_FigS1, breaks_vector_n = breaks_vector_n) 
zoo_plot_emp2_S1B<-zoo_plot_emp2_S1B+ggtitle("b) Zooplankton empirical")

### ESMs data for Fig S1B -----

phyto_plot2_S1B<-plot_SO(wmap_df = antmap_df, data_to_plot = mean_phyto, proj = circumpolarCRS, legend_name = legend_name, min =mean_phyto@data@min, max = mean_phyto@data@max, data_type = "modelled", color_scale = color_scale, theme_map= "my_theme_map2", scale = scale_FigS1, breaks_vector_n = breaks_vector_n)
phyto_plot2_S1B<-phyto_plot2_S1B+ggtitle("c) Phytoplankton modelled")

zoo_plot2_S1B<-plot_SO(wmap_df = antmap_df, data_to_plot = mean_zoo, proj = circumpolarCRS, legend_name = legend_name, min =mean_zoo@data@min, max = mean_zoo@data@max, data_type = "modelled", color_scale = color_scale, theme_map= "my_theme_map2", scale = scale_FigS1, breaks_vector_n = breaks_vector_n) 
zoo_plot2_S1B<-zoo_plot2_S1B+ggtitle("d) Zooplankton modelled")

### MEMs data for Fig S2 -----

# Fig S2. Depth integrated, size-resolved total consumers biomass averaged across the 1990-1999 decade and six FishMIP Marine Ecosystem Models able to produce this output. Size classes include modelled individuals of a) 1-10g, b) 10-100g, c) 100g-1kg, d) 1-10kg, e) 10-100kg, f) 100kg-1t. All Marine Ecosystem Models are forced using IPSL-CM6A-LR and GFDL-ESM4 climate data with the exception of two models which are forced using IPSL-CM6A-LR climate data only, for a total of ten ESMs-MEMs combinations. The simulation round considered is the ISIMIP3b, data available at https://data.isimip.org/. The range of values is kept between 0.5 and 5 for all size classes for better comparison across plots and with maps of plankton (S-Fig.1). The code used to produce this figure is available at https://github.com/Fish-MIP/Extract_SouthernOcean.

if(scale_FigS2 == "raw"){
  legend_name = expression(g~C~m^{-2})
}else if(scale_FigS2 == "log"){
  legend_name = expression(log~g~C~m^{-2}) 
}

size1_plot<-plot_SO(wmap_df = antmap_df, data_to_plot = mean_all[[1]], proj = circumpolarCRS, legend_name = legend_name, min = min_FigS2, max = max_FigS2, data_type = "modelled", color_scale = color_scale, theme_map= "my_theme_map2", scale = scale_FigS2, breaks_vector_n = breaks_vector_n)
size1_plot<-size1_plot+ggtitle("a) Consumers Biomass 1-10g")
# plot(mean_all[[1]])

size2_plot<-plot_SO(wmap_df = antmap_df, data_to_plot = mean_all[[2]], proj = circumpolarCRS, legend_name = legend_name, min = min_FigS2, max = max_FigS2, data_type = "modelled", color_scale = color_scale, theme_map= "my_theme_map2", scale = scale_FigS2, breaks_vector_n = breaks_vector_n)
size2_plot<-size2_plot+ggtitle("b) Consumers Biomass 10-100g")
# plot(mean_all[[2]])

size3_plot<-plot_SO(wmap_df = antmap_df, data_to_plot = mean_all[[3]], proj = circumpolarCRS, legend_name = legend_name, min = min_FigS2, max = max_FigS2, data_type = "modelled", color_scale = color_scale, theme_map= "my_theme_map2", scale = scale_FigS2, breaks_vector_n = breaks_vector_n)
size3_plot<-size3_plot+ggtitle("c) Consumers Biomass 100g-1kg")
# plot(mean_all[[3]])

size4_plot<-plot_SO(wmap_df = antmap_df, data_to_plot = mean_all[[4]], proj = circumpolarCRS, legend_name = legend_name, min = min_FigS2, max = max_FigS2, data_type = "modelled", color_scale = color_scale, theme_map= "my_theme_map2", scale = scale_FigS2, breaks_vector_n = breaks_vector_n)
size4_plot<-size4_plot+ggtitle("d) Consumers Biomass 1-10kg")
# plot(mean_all[[4]])

size5_plot<-plot_SO(wmap_df = antmap_df, data_to_plot = mean_all[[5]], proj = circumpolarCRS, legend_name = legend_name, min = min_FigS2, max = max_FigS2, data_type = "modelled", color_scale = color_scale, theme_map= "my_theme_map2", scale = scale_FigS2, breaks_vector_n = breaks_vector_n)
size5_plot<-size5_plot+ggtitle("e) Consumers Biomass 10-100kg")
# plot(mean_all[[5]])

size6_plot<-plot_SO(wmap_df = antmap_df, data_to_plot = mean_all[[6]], proj = circumpolarCRS, legend_name = legend_name, min = min_FigS2, max = max_FigS2, data_type = "modelled", color_scale = color_scale, theme_map= "my_theme_map2", scale = scale_FigS2, breaks_vector_n = breaks_vector_n)
size6_plot<-size6_plot+ggtitle("f) Consumers Biomass >100kg")
# plot(mean_all[[6]])


### ESMs and MEMs data for FigS3 ----

# Fig S3. Trophic amplification - ... TO ADJUST: Depth integrated, size-resolved total consumers biomass averaged across the 1990-1999 decade and six FishMIP Marine Ecosystem Models able to produce this output. Size classes include modelled individuals of a) 1-10g, b) 10-100g, c) 100g-1kg, d) 1-10kg, e) 10-100kg, f) 100kg-1t. All Marine Ecosystem Models are forced using IPSL-CM6A-LR and GFDL-ESM4 climate data with the exception of two models which are forced using IPSL-CM6A-LR climate data only, for a total of ten ESMs-MEMs combinations. The simulation round considered is the ISIMIP3b, data available at https://data.isimip.org/. The range of values is kept between 0.5 and 5 for all size classes for better comparison across plots and with maps of plankton (S-Fig.1). The code used to produce this figure is available at https://github.com/Fish-MIP/Extract_SouthernOcean.

if(scale_FigS3 == "raw"){
  legend_name = expression(g~C~m^{-2})
}else if(scale_FigS3 == "log"){
  legend_name = expression(log~g~C~m^{-2}) 
}

# names(ltl)<-"index_1" 
# ltl_plot<-plot_SO(wmap_df = antmap_df, data_to_plot = ltl, proj = circumpolarCRS, legend_name = legend_name, min = ltl@data@min, max = ltl@data@max, data_type = "modelled", color_scale = color_scale, theme_map= "my_theme_map2", scale = scale_FigS3, breaks_vector_n = breaks_vector_n)
# 
# if(ltl_type == "phyto"){
#  ltl_plot<-ltl_plot+ggtitle("a) Lower trophic level (phyto)") 
# }else if(ltl_type == "zoop"){
#   ltl_plot<-ltl_plot+ggtitle("a) Lower trophic level (zoop)")
# }else if(ltl_type == "sum"){
#   ltl_plot<-ltl_plot+ggtitle("a) Lower trophic level (phyto + zoop)")
# }

color_scale = "divergent"
breaks_vector_n = 11


names(tl2)<-"index_1"
tl2_plot<-plot_SO(wmap_df = antmap_df, data_to_plot = tl2, proj = circumpolarCRS, legend_name = legend_name, min = min_FigS3, max = max_FigS3, data_type = "modelled", color_scale = color_scale, theme_map= "my_theme_map2", scale = scale_FigS3, breaks_vector_n = breaks_vector_n) 

if(ltl_type == "phyto"){
 tl2_plot<-tl2_plot+ggtitle("a) Biomass 1-10g / phytoplankton")
}else if(ltl_type == "zoop"){
  tl2_plot<-tl2_plot+ggtitle("a) Biomass 1-10g / zooplankton")
}else if(ltl_type == "sum"){
  tl2_plot<-tl2_plot+ggtitle("a) Biomass 1-10g / LTL")
}

names(tl3)<-"index_1"
tl3_plot<-plot_SO(wmap_df = antmap_df, data_to_plot = tl3, proj = circumpolarCRS, legend_name = legend_name, min = min_FigS3, max = max_FigS3, data_type = "modelled", color_scale = color_scale, theme_map= "my_theme_map2", scale = scale_FigS3, breaks_vector_n = breaks_vector_n)
tl3_plot<-tl3_plot+ggtitle("b) Biomass 10-100g / 1-10g")

names(tl4)<-"index_1"
tl4_plot<-plot_SO(wmap_df = antmap_df, data_to_plot = tl4, proj = circumpolarCRS, legend_name = legend_name, min = min_FigS3, max = max_FigS3, data_type = "modelled", color_scale = color_scale, theme_map= "my_theme_map2", scale = scale_FigS3, breaks_vector_n = breaks_vector_n)
tl4_plot<-tl4_plot+ggtitle("c) Biomass 100g-1kg / 10-100g")

names(tl5)<-"index_1"
tl5_plot<-plot_SO(wmap_df = antmap_df, data_to_plot = tl5, proj = circumpolarCRS, legend_name = legend_name, min = min_FigS3, max = max_FigS3, data_type = "modelled", color_scale = color_scale, theme_map= "my_theme_map2", scale = scale_FigS3, breaks_vector_n = breaks_vector_n)
tl5_plot<-tl5_plot+ggtitle("d) Biomass 1-10kg / 100g-1kg")

names(tl6)<-"index_1"
tl6_plot<-plot_SO(wmap_df = antmap_df, data_to_plot = tl6, proj = circumpolarCRS, legend_name = legend_name, min = min_FigS3, max = max_FigS3, data_type = "modelled", color_scale = color_scale, theme_map= "my_theme_map2", scale = scale_FigS3, breaks_vector_n = breaks_vector_n)
tl6_plot<-tl6_plot+ggtitle("e) Biomass 10-100kg / 1-10kg")

names(tl7)<-"index_1"
tl7_plot<-plot_SO(wmap_df = antmap_df, data_to_plot = tl7, proj = circumpolarCRS, legend_name = legend_name, min = min_FigS3, max = max_FigS3, data_type = "modelled", color_scale = color_scale, theme_map= "my_theme_map2", scale = scale_FigS3, breaks_vector_n = breaks_vector_n)
tl7_plot<-tl7_plot+ggtitle("f) Biomass >100kg / 10-100kg")

```

## Save fianl plots 

```{r}

#### Fig 0 check empirical ----

p_final_Fig0<-
  check_phyto + check_zoo +
  plot_layout(ncol = 2, nrow = 1)

name_Fig0 = paste("/home/ubuntu/Extract_SouthernOcean/Output/", Sys.Date(),"_" ,"Fig0_checkEmpirical","_",scale_Fig0,".pdf", sep = "")

pdf(name_Fig0, width = 7, height = 4.4, bg = "transparent")  
p_final_Fig0
dev.off()

#### Fig 5 ----
layout <- "
AB
CD
"

p_final_Fig4<-
  phyto_emp_plot + zoop_emp_plot +
  phyto_plot + zoo_plot +
  # plot_layout(design = layout)
  plot_layout(ncol = 2, nrow = 2, guides = 'collect', heights=c(1,1), widths = c(1,1))

name_Fig4 = paste("/home/ubuntu/Extract_SouthernOcean/Output/", Sys.Date(),"_" ,"Fig4","_",scale_Fig4,"_",removeNA,".pdf", sep = "")

# name_Fig4<-"/home/ubuntu/Extract_SouthernOcean/Output/trial.pdf"

# changing width and height fix the problem but it needs to be changed by a lot 10X10
pdf(name_Fig4, width = 6, height = 6, bg = "transparent") 
p_final_Fig4
dev.off()

# # when too small ... words cut off 
# pdf("/home/ubuntu/Extract_SouthernOcean/Output/trial.pdf", width = 3.5, height = 3.5, bg = "transparent") 
# zoop_emp_plot
# dev.off()


#### Fig S1 ----

p_final_FigS1<-
  phyto_plot_emp2 + zoo_plot_emp2 +
  phyto_plot2 + zoo_plot2 +
  plot_layout(ncol = 2, nrow = 2, heights=c(1,1,1,1), widths = c(1,1,1,1), guides = 'collect')

name_FigS1 = paste("/home/ubuntu/Extract_SouthernOcean/Output/", Sys.Date(),"_" ,"FigS1","_",scale_FigS1,"_",removeNA,".pdf", sep = "")

pdf(name_FigS1, width = 6, height = 6, bg = "transparent")
p_final_FigS1
dev.off()

#### Fig S1B ----

p_final_FigS1B<-
  phyto_plot_emp2_S1B + zoo_plot_emp2_S1B +
  phyto_plot2_S1B + zoo_plot2_S1B +
  plot_layout(ncol = 2, nrow = 2, heights=c(1,1,1,1), widths = c(1,1,1,1)) # guides = 'collect'

name_FigS1B = paste("/home/ubuntu/Extract_SouthernOcean/Output/", Sys.Date(),"_" ,"FigS1B","_",scale_FigS1,"_",removeNA,".pdf", sep = "")

pdf(name_FigS1B, width = 6, height = 6, bg = "transparent")
p_final_FigS1B
dev.off()

#### Fig S2 ----

p_final_FigS2<-size1_plot + size2_plot +
  size3_plot + size4_plot +
  size5_plot + size6_plot +
  plot_layout(ncol = 3, nrow = 2, guides = 'collect')

name_FigS2 = paste("/home/ubuntu/Extract_SouthernOcean/Output/", Sys.Date(),"_" ,"FigS2","_",scale_FigS2,"_",removeNA,".pdf", sep = "")

pdf(name_FigS2, width = 8, height = 7, bg = "transparent")
p_final_FigS2
dev.off()

#### Fig S3 ----

p_final_FigS3<-tl2_plot + tl3_plot +
  tl4_plot + tl5_plot +
  tl6_plot + tl7_plot +
  plot_layout(ncol = 3, nrow = 2, guides = 'collect')

name_FigS3 = paste("/home/ubuntu/Extract_SouthernOcean/Output/", Sys.Date(),"_" ,"FigS3","_",scale_FigS2,"_",removeNA,"_ltl_type_",ltl_type,".pdf", sep = "")

pdf(name_FigS3, width = 8, height = 7, bg = "transparent")
p_final_FigS3
dev.off()


```


